$K$-corrections and photometric redshifts in SDSS and 2MASS$^1$

### 2004-01-17 docs/paper/kcorrect.tex
add
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@292 75dff457-8751-0410-ab1a-cd701654ce6b

%% You can insert a short comment on the title page using the command below.

@@

%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
%% The left head contains a list of authors,
%% usually a maximum of three (otherwise use et al.).  The right
%% head is a modified title of up to roughly 44 characters.  Running heads
%% will not print in the manuscript style.

\shortauthors{Blanton {\it et al.} (2000)}
\shorttitle{Modeling Galaxy SEDs}

%% This paper uses runs 752/756, rerun 4
%% It uses redshifts from version 3 of spectro1d

%% This is the end of the preamble.  Indicate the beginning of the
%% paper itself with \begin{document}.

@@

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire.

\title{Estimating Fixed Frame Galaxy Magnitudes in the SDSS$^1$}

%% Use \author, \affil, and the \and command to format
%% author and affiliation information.
%% Note that \email has replaced the old \authoremail command
%% from AASTeX v4.0. You can use \email to mark an email address
%% anywhere in the paper, not just in the front matter.
%% As in the title, you can use \\ to force line breaks.

%\author{Michael Blanton}
%\affil{NASA/Fermilab Astrophysics Center\\
%Fermi National Accelerator Laboratory, Batavia, IL 60510-0500}
%\author{\and lots and lots of other, probably more important people}
%\affil{Some Institution\\
%Somewhere, Some City, Some State or Province}
%\email{blanton@fnal.gov}

% Authorship determined by those I was directly involved with
% in performing this work as well as those responsible for the photo-z
% plates and those who have characterized the filter curves (since
% these might be publicly distributed). 
\author{
Michael R. Blanton\altaffilmark{\ref{NYU}},
%Tamas Budavari\altaffilmark{\ref{JHU}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
J.~Brinkmann\altaffilmark{\ref{APO}},
Istv\'an Csabai\altaffilmark{\ref{JHU}},
Mamoru Doi\altaffilmark{\ref{Tokyo}},
Daniel Eisenstein\altaffilmark{\ref{Arizona}},
Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
James E. Gunn\altaffilmark{\ref{Princeton}},
David W. Hogg\altaffilmark{\ref{NYU}}, and
David J. Schlegel\altaffilmark{\ref{Princeton}}
%Julianne Dalcanton\altaffilmark{\ref{UW}},
%Jon Loveday\altaffilmark{\ref{Sussex}},
%Michael A. Strauss\altaffilmark{\ref{Princeton}},
%Mark SubbaRao\altaffilmark{\ref{Chicago}},
%David H. Weinberg\altaffilmark{\ref{Ohio}},
%John E. Anderson, Jr.\altaffilmark{\ref{Fermilab}},
%James Annis\altaffilmark{\ref{Fermilab}},
%Neta A. Bahcall\altaffilmark{\ref{Princeton}},
%Mariangela Bernardi\altaffilmark{\ref{Chicago}},
%Robert J. Brunner\altaffilmark{\ref{Caltech}},
%Scott Burles\altaffilmark{\ref{Fermilab}},
%Larry Carey\altaffilmark{\ref{UW}},
%Francisco J. Castander\altaffilmark{\ref{Chicago}, \ref{Pyrenees}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Douglas Finkbeiner\altaffilmark{\ref{Berkeley}},
%Scott Friedman\altaffilmark{\ref{JHU}},
%Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
%G. S. Hennessy\altaffilmark{\ref{USNO}},
%Robert B. Hindsley\altaffilmark{\ref{USNO}},
%Takashi Ichikawa\altaffilmark{\ref{Tokyo}},
%\v{Z}eljko Ivezi\'{c}\altaffilmark{\ref{Princeton}},
%Stephen Kent\altaffilmark{\ref{Fermilab}},
%G. R.~Knapp\altaffilmark{\ref{Princeton}},
%D. Q.~Lamb\altaffilmark{\ref{Chicago}},
%R. French Leger\altaffilmark{\ref{UW}},
%Daniel C. Long\altaffilmark{\ref{APO}},
%Robert H. Lupton\altaffilmark{\ref{Princeton}},
%Timothy A.~McKay\altaffilmark{\ref{Michigan}},
%Avery Meiksin\altaffilmark{\ref{Edinburgh}},
%Aronne Merelli\altaffilmark{\ref{Caltech}},
%Jeffrey A. Munn\altaffilmark{\ref{USNO}},
%Vijay Narayanan\altaffilmark{\ref{Princeton}},
%Matt Newcomb\altaffilmark{\ref{CarnegieMellon}},
%R. C. Nichol\altaffilmark{\ref{CarnegieMellon}},
%Sadanori Okamura\altaffilmark{\ref{Tokyo}},
%Russell Owen\altaffilmark{\ref{UW}},
%Jeffrey R.~Pier\altaffilmark{\ref{USNO}},
%Adrian Pope\altaffilmark{\ref{JHU}},
%Marc Postman\altaffilmark{\ref{STScI}},
%Thomas Quinn\altaffilmark{\ref{UW}},
%Constance M. Rockosi\altaffilmark{\ref{Chicago}},
%Donald P. Schneider\altaffilmark{\ref{PennState}}, 
%Kazuhiro Shimasaku\altaffilmark{\ref{Tokyo}},
%Walter A. Siegmund\altaffilmark{\ref{UW}},
%Stephen Smee\altaffilmark{\ref{Maryland}},
%Yehuda Snir\altaffilmark{\ref{CarnegieMellon}},
%Chris Stoughton\altaffilmark{\ref{Fermilab}},
%Christopher Stubbs\altaffilmark{\ref{UW}},
%Alexander S.~Szalay\altaffilmark{\ref{JHU}},
%Gyula P.~Szokoly\altaffilmark{\ref{Potsdam}},
%Aniruddha R.~Thakar\altaffilmark{\ref{JHU}},
%Christy Tremonti\altaffilmark{\ref{JHU}},
%Douglas L. Tucker\altaffilmark{\ref{Fermilab}},
%Alan Uomoto\altaffilmark{\ref{JHU}},
%Dan vanden Berk\altaffilmark{\ref{Fermilab}},
%Michael S. Vogeley\altaffilmark{\ref{Drexel}},
%Patrick Waddell\altaffilmark{\ref{UW}},
%Brian Yanny\altaffilmark{\ref{Fermilab}},
%Naoki Yasuda\altaffilmark{\ref{NAOJ}},
%and Donald G.~York\altaffilmark{\ref{Chicago}}
}

@@

Department of Physics and Astronomy, The Johns Hopkins University,
Baltimore, MD 21218
\label{JHU}}

@@

Department of Astronomy and Research Center for 
the Early Universe,
School of Science, University of Tokyo,
Tokyo 113-0033, Japan
\label{Tokyo}}
Steward Observatory, 
933 N. Cherry Ave., Tucson, AZ
85721
\label{Arizona}}
Princeton University Observatory, Princeton,
NJ 08544
\label{Princeton}}

@@

Apache Point Observatory,
2001 Apache Point Road,
P.O. Box 59, Sunspot, NM 88349-0059
\label{APO}}

@@

Institute for Cosmic Ray Research, University of
Tokyo, Midori, Tanashi, Tokyo 188-8502, Japan
\label{CosmicRay}}
Institute for Advanced Study, Olden Lane,
Princeton, NJ 08540
\label{IAS}}

@@

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.
\begin{abstract}
Broad-band measurements of flux for galaxies at different redshifts
measure different regions of the rest-frame galaxy spectrum. Certain
astronomical questions, such as the evolution of the luminosity
function of galaxies, require transforming these inherently
redshift-dependent magnitudes into redshift-independent quantities. To
prepare to address these astronomical questions, investigated in
detail in subsequent papers, we fit spectral energy distributions
(SEDs) to broad band photometric observations, in the context of the
optical observations of the Sloan Digital Sky Survey (SDSS). Linear
combinations of four spectral templates can reproduce the five SDSS
magnitudes of all galaxies to the precision of the
photometry. Expressed in the appropriate coordinate system, the
locus of the coefficients multiplying the templates is planar,
and in fact nearly linear. The resulting reconstructed SEDs can be
used to recover fixed frame magnitudes over a range of redshifts. This
process yields consistent results, in the sense that within each
sample the intrinsic colors of similar type galaxies are nearly
constant with redshift. We compare our results to simpler
interpolation methods and galaxy spectrophotometry from the SDSS.  The
software that generates these results is publicly available and easily
adapted to handle a wide range of galaxy observations.
\end{abstract}

@@

%
% Introduction and motivation
%

@@

In the future, observations of galaxies (and indeed of any
astronomical sources) will be performed using devices which combine
high spatial resolution and high spectral resolution. By that time,
the phenomena of broad band filters and the quaint terminology
surrounding their usage --- magnitudes, $K$-corrections, color terms,
{\it etc.} --- will have long since been forgotten. However, until
such time as wide field spatially-resolved spectroscopy is cheaply
available, virtually all observations of galaxies will be made through
broad band filters, and special care has to be taken to recover
knowledge of galaxy spectral energy distributions (SEDs) from these
observations.  Reconstructing galaxy SEDs is nontrivial because SEDs
of galaxies contain important information on scales considerably
smaller than the width of typical broad band filters. Furthermore,
SEDs of distant galaxies are redshifted such that the more distant the
galaxy, the further the observed bandpass is blueshifted relative to
the rest-frame spectral energy distribution of the object observed.
This paper focuses on a solution to these problems which (in the
optical wavelength regime) is fast, robust, and consistent over a
large range of redshifts.

In the past, people have accounted for these effects using
``$K$-corrections'' applied to the observed magnitudes. In these
analyses, a function $K(z)$ is added to the standard cosmological
bolometric distance modulus $\mathrm{DM}(z)$ to obtain the
relationship between the apparent magnitude $m_R$ of band $R$ and the
absolute magnitude $M$ of (in general different) band $Q$:
\begin{equation}
\label{kcorrecteqn}
m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z)
\end{equation}
The traditional definition of the $K$-correction takes $Q=R$; however,
we note that in practice many surveys do perform $K$-corrections from
one bandpass to another when comparing high redshift and low redshift
observations.

Sometimes a single function $K(z)$ has been applied regardless of the
galaxy type, though more recently it has become standard to use a
discrete set of $K(z)$ functions depending on galaxy type (based
either on morphology or spectral features). As described in
\cite{oke68a} and later papers, the function is based on the
projection of an assumed galaxy SED redshifted to $z$ onto the
measured instrument response as a function of wavelength (including
the effects of the atmosphere, the reflectivity of the mirrors, the
filter transmission, and the response of the CCD device or
photographic emulsion). For example, for the Sloan Digital Sky Survey
(SDSS) filter system, \citet{frei94a} and \citet{fukugita95a} have
both presented results for the $K$-corrections of galaxies. Each of
these papers tabulates $K(z)$ (called $k(z)$ by \citealt{frei94a}) for
galaxies of various morphological types, based on galaxy
spectrophotometry.\footnote{Note that the description on p. 57 of
\citet{binney98a} of the meaning of $k(z)$ in \citet{frei94a} is
incorrect; in fact, their $k(z)$ is exactly equivalent to our $K(z)$.}

However, galaxies do not all have the same SED, nor are they selected
from some discrete set of SEDs. For this reason, these schemes for
applying the $K$-corrections can fail to be self-consistent: the SED
assumed for the $K$-corrections can be significantly inconsistent with
the observed galaxy colors! As we demand more precision from our
astronomical data analysis in new, large multi-band surveys, and in
particular as we try to quantify the evolution of galaxies, we must
take a more sophisticated approach to approximating fixed frame
observations of galaxies.

Our approach here is to use a method for inferring the underlying SEDs
of a set of galaxies at a range of redshifts by requiring that their
SEDs all be drawn from a similar population. For each galaxy, we will
recover a model SED, which can be used to synthesize the galaxy's
magnitude in any bandpass. Although the operation we are performing on
the magnitudes is not a ``$K$-correction'' in the historical sense of
the term, we will refer to it as such in this and subsequent
papers. Our approach is equivalent (nearly identical) to the
photometric redshift estimation methods of \citet{csabai00a},
except we use slightly different coordinate systems for our SED
template space. In fact, the software includes a fast and relatively
accurate photometric redshift estimator based on their method.

We implemented this system and are publishing it in order to lay the
groundwork for upcoming papers which will rely heavily on the
reliability of the fixed frame magnitudes determined here.  Because
the SDSS is one of the largest astronomical surveys to date, our
method will be useful to a large number of other investigators, both
inside and outside the SDSS collaboration.  Thus, this paper also
serves the purpose of describing the release of a piece of
software. The computer software we distribute builds itself into a C
shared object library, around which we have written both stand-alone C
programs and IDL routines. We have made the source code available
publicly, through a web page and through a public CVS
repository. Improvements or ports to other languages implemented by
users may be incorporated into the code upon request. The conditions
of use for the code are that this paper be cited in any resulting
refereed journal article and that the version of the code used is
specified in any such paper. The version of the code used to make the
figures for this paper is {\tt kcorrect \kversion}.

We refer throughout this paper to $AB$ magnitudes, first defined by
\citet{oke83a} to measure the ratio of the number of photons included
in the signal of the detector relative to that number for a flat
spectrum source with $g(\nu)=3.631 \times 10^{-20}$ ergs cm$^{-2}$
s$^{-1}$ Hz$^{-1}$. For a source with a spectrum $f(\nu)$ the $AB$
magnitude should be (for a perfectly calibrated $AB$ system)
\begin{eqnarray}
\label{ABdef}
m_{AB} &=& -2.41 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
\cr
&=& -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
\end{eqnarray}
where $R(\lambda)$ is the fraction of photons entering the Earth's
atmosphere which are included in the signal as a function of
wavelength (a unitless quantity). Note that $R(\lambda)$ can be
defined even for devices which do not count photons directly (such as
bolometers). This equation is written such that $f(\lambda)$ is in
units of ergs cm$^{-2}$ s$^{-1}$ \AA$^{-1}$ and $f(\nu)$ is in units
of ergs cm$^{-2}$ s$^{-1}$ Hz$^{-1}$, while $\lambda$ is expressed in
\AA\ and $\nu$ is expressed in Hz. The normalizations defined here
mean that an object with $f(\nu) = g(\nu) = 3631\mathrm{~Jy} = 3.631
\times 10^{-20}$ ergs cm$^{-2}$ s$^{-1}$ Hz$^{-1}$ has all its AB
magnitudes equal to zero.  The $\lambda^{-1}$ appears in the integrand
of the denominator of the first equation because
$g(\lambda)=c/\lambda^2$ for a ``flat spectrum'' source with
$g(\nu)=1$.  The difference in the zeropoints of the two equations
simply corresponds to the factor of the speed of light $c$ (expressed
in \AA\ s$^{-1}$) in that expression for $g(\lambda)$.

Section \ref{sedfit} describes our method of fitting galaxy SEDs to
broad band photometry and how to calculate $K$-corrections. Section
\ref{data} applies the method to galaxies in the SDSS, showing that
the fits are robust.  Section \ref{conclusions} concludes and
discusses future development of the method described here.

\section{Calculating Fixed Frame Galaxy Magnitudes}
\label{sedfit}

First, we describe how we reconstruct galaxy SEDs from broad band
magnitude measurements. Second, we describe how we convert these SEDs
into estimates of fixed frame galaxy magnitudes.

\subsection{Fitting SEDs to Galaxy Broad-band Magnitudes}

Our task is to recover a model for the galaxy SED from broad band
photometric measurements. Since a set of broad band galaxy fluxes does
not correspond uniquely to a particular SED, and because galaxy SEDs
are known to have significant structure over wavelength ranges small
compared to our bandpass, the inverse problem of reconstructing SEDs
from broad band fluxes is ill-posed. The task is made yet more
difficult by the fact that the separation of the filters are large
compared to their widths; the SEDs are thus not well-sampled in the
Nyquist sense of that term. However, we are not completely ignorant
about the forms which galaxy SEDs take, so we can attempt to use what
we know about galaxy SEDs to appropriately regularize our fits.  The
method described here for doing so follows closely the method of
estimating photometric redshifts used by \citet{csabai00a}.  It is
designed to take advantage of what we already understand about galaxy
spectral energy distributions.
%In addition, as implemented here, the method is
%equivalent to a fit to the star-formation history of the galaxies
%using a spectrophotometric model; we will investigate the implications
%of the model fits in a separate paper and concentrate here on the
%empirical aspects of the results.

Begin with an SED space defined by $N_b$ template galaxy SEDs
$\vv{v}_i(\lambda)$ (for example from \citealt{bruzual93a}), where $N_b$ is
large. In principle, this should be a complete set of galaxy SEDs if
we want {\it all} galaxy SEDs to be in the space spanned by the
$\vv{v}_i(\lambda)$; in practice, this property is not necessary.
Given a set of $\vv{v}_i(\lambda)$, one can find an orthonormal set of
spectra $\vv{b}_i(\lambda)$ spanning the same space.  To be more
precise, we define the dot-product in the space of galaxy spectra to
be:
\begin{equation}
\vv{v}_i \cdot \vv{v}_j =
\int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
v_i(\lambda) v_j(\lambda),
\end{equation}
where $\lambda_{\mathrm{min}}$ and $\lambda_{\mathrm{max}}$ define the
wavelength range used to define orthogonality. These limits are
defined not over the full range covered by the $\vv{v}_i(\lambda)$
spectra, but instead over a smaller range which is within observable
wavelengths over most of the redshift range of the sample.\footnote{
If the template spectra were identical in the subrange chosen and only
differed outside of it, our choice would cause computational problems,
but in fact the spectra are all independent in the restricted, as well
as the full, wavelength range.}  We then define the basis of the SED
space such that the $N_b$ template SEDs, $\vv{v}_i$, are linear
combinations of $N_b$ basis spectra $\vv{b}_i$ and that $\vv{b}_i
\cdot \vv{b}_j = \delta_{ij}$ (where $\delta_{ij}$ represents the
Kronecker delta). These conditions do not fully specify the
$\vv{b}_i$, which naturally may be rotated arbitrarily within our
$N_b$-dimensional subspace of spectral space.

>>>>

@@

\shortauthors{Blanton {\it et al.} (2000)}
\shorttitle{$K$-corrections and photometric redshifts}

% TODO:
% finish final templates
% fix kphotoz banding
% fit SFHs to final spectra to get stellar masses, mean ages, 
% mean metallicities

@@

\title{$K$-corrections and photometric redshifts in SDSS and 2MASS$^1$}

\author{
Michael R. Blanton\altaffilmark{\ref{NYU}}
%Tamas Budavari\altaffilmark{\ref{JHU}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%J.~Brinkmann\altaffilmark{\ref{APO}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Mamoru Doi\altaffilmark{\ref{Tokyo}},
%Daniel Eisenstein\altaffilmark{\ref{Arizona}},
%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
%James E. Gunn\altaffilmark{\ref{Princeton}},
%David W. Hogg\altaffilmark{\ref{NYU}}, and
%David J. Schlegel\altaffilmark{\ref{Princeton}}
%Julianne Dalcanton\altaffilmark{\ref{UW}},
%Jon Loveday\altaffilmark{\ref{Sussex}},
%Michael A. Strauss\altaffilmark{\ref{Princeton}},
%Mark SubbaRao\altaffilmark{\ref{Chicago}},
%David H. Weinberg\altaffilmark{\ref{Ohio}},
%John E. Anderson, Jr.\altaffilmark{\ref{Fermilab}},
%James Annis\altaffilmark{\ref{Fermilab}},
%Neta A. Bahcall\altaffilmark{\ref{Princeton}},
%Mariangela Bernardi\altaffilmark{\ref{Chicago}},
%Robert J. Brunner\altaffilmark{\ref{Caltech}},
%Scott Burles\altaffilmark{\ref{Fermilab}},
%Larry Carey\altaffilmark{\ref{UW}},
%Francisco J. Castander\altaffilmark{\ref{Chicago}, \ref{Pyrenees}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Douglas Finkbeiner\altaffilmark{\ref{Berkeley}},
%Scott Friedman\altaffilmark{\ref{JHU}},
%Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
%G. S. Hennessy\altaffilmark{\ref{USNO}},
%Robert B. Hindsley\altaffilmark{\ref{USNO}},
%Takashi Ichikawa\altaffilmark{\ref{Tokyo}},
%\v{Z}eljko Ivezi\'{c}\altaffilmark{\ref{Princeton}},
%Stephen Kent\altaffilmark{\ref{Fermilab}},
%G. R.~Knapp\altaffilmark{\ref{Princeton}},
%D. Q.~Lamb\altaffilmark{\ref{Chicago}},
%R. French Leger\altaffilmark{\ref{UW}},
%Daniel C. Long\altaffilmark{\ref{APO}},
%Robert H. Lupton\altaffilmark{\ref{Princeton}},
%Timothy A.~McKay\altaffilmark{\ref{Michigan}},
%Avery Meiksin\altaffilmark{\ref{Edinburgh}},
%Aronne Merelli\altaffilmark{\ref{Caltech}},
%Jeffrey A. Munn\altaffilmark{\ref{USNO}},
%Vijay Narayanan\altaffilmark{\ref{Princeton}},
%Matt Newcomb\altaffilmark{\ref{CarnegieMellon}},
%R. C. Nichol\altaffilmark{\ref{CarnegieMellon}},
%Sadanori Okamura\altaffilmark{\ref{Tokyo}},
%Russell Owen\altaffilmark{\ref{UW}},
%Jeffrey R.~Pier\altaffilmark{\ref{USNO}},
%Adrian Pope\altaffilmark{\ref{JHU}},
%Marc Postman\altaffilmark{\ref{STScI}},
%Thomas Quinn\altaffilmark{\ref{UW}},
%Constance M. Rockosi\altaffilmark{\ref{Chicago}},
%Donald P. Schneider\altaffilmark{\ref{PennState}}, 
%Kazuhiro Shimasaku\altaffilmark{\ref{Tokyo}},
%Walter A. Siegmund\altaffilmark{\ref{UW}},
%Stephen Smee\altaffilmark{\ref{Maryland}},
%Yehuda Snir\altaffilmark{\ref{CarnegieMellon}},
%Chris Stoughton\altaffilmark{\ref{Fermilab}},
%Christopher Stubbs\altaffilmark{\ref{UW}},
%Alexander S.~Szalay\altaffilmark{\ref{JHU}},
%Gyula P.~Szokoly\altaffilmark{\ref{Potsdam}},
%Aniruddha R.~Thakar\altaffilmark{\ref{JHU}},
%Christy Tremonti\altaffilmark{\ref{JHU}},
%Douglas L. Tucker\altaffilmark{\ref{Fermilab}},
%Alan Uomoto\altaffilmark{\ref{JHU}},
%Dan vanden Berk\altaffilmark{\ref{Fermilab}},
%Michael S. Vogeley\altaffilmark{\ref{Drexel}},
%Patrick Waddell\altaffilmark{\ref{UW}},
%Brian Yanny\altaffilmark{\ref{Fermilab}},
%Naoki Yasuda\altaffilmark{\ref{NAOJ}},
%and Donald G.~York\altaffilmark{\ref{Chicago}}
}

@@

%Department of Physics and Astronomy, The Johns Hopkins University,
%Baltimore, MD 21218
%\label{JHU}}

@@

%Department of Astronomy and Research Center for 
%the Early Universe,
%School of Science, University of Tokyo,
%Tokyo 113-0033, Japan
%\label{Tokyo}}
%Steward Observatory, 
%933 N. Cherry Ave., Tucson, AZ
%85721
%\label{Arizona}}
%Princeton University Observatory, Princeton,
%NJ 08544
%\label{Princeton}}

@@

%Apache Point Observatory,
%2001 Apache Point Road,
%P.O. Box 59, Sunspot, NM 88349-0059
%\label{APO}}

@@

%Institute for Cosmic Ray Research, University of
%Tokyo, Midori, Tanashi, Tokyo 188-8502, Japan
%\label{CosmicRay}}
%Institute for Advanced Study, Olden Lane,
%Princeton, NJ 08540
%\label{IAS}}

@@

\begin{abstract}
We present a method to develop templates for calculating
$K$-corrections and photometric redshifts suitable for ultraviolet,
optical, and infrared observations. Since we base the templates on
stellar population synthesis results, the results are intepretable in
terms of approximate stellar masses and star-formation histories. We
present templates fit with this method to the Sloan Digital Sky Survey
and Two-Micron All Sky Survey photometry.  The software that generates
these results is publicly available and easily adapted to handle a
wide range of galaxy observations.
\end{abstract}

@@

@@

\section{Fitting SEDs to broad-band magnitudes}
\label{sedfit}

First, we describe how we reconstruct galaxy SEDs from broad band
magnitude measurements.

A broad-band flux $F_k$ in filter $k$ is a measurement of the
projection of the source spectrum $f(\lambda)$ on the response
$R_k(\lambda)$ (where typically this function includes the effects of
the atmosphere, telescope, filter $k$, and the CCD itself)
\begin{equation}
F_k = \int_0^\infty d\lambda f(\lambda) R_k(\lambda).
\end{equation}
Now, let us model $f(\lambda)$ as a sum of many template spectra:
\begin{equation}
f(\lambda) = \sum_j a_j v_j(\lambda)
\end{equation}
In order to fit the $a_j$ given the observations in the bands $k$ for
a single object, we minimize
\begin{eqnarray}
\label{chi2}
\chi^2 &=& \sum_{k} \left[ \frac{F_{k,\mathrm{obs}} - F(k)}{\sigma_k}
\right]^2 \cr
&=& \sum_{ik} \left[ \frac{F_{k,\mathrm{obs}} -\sum_j a_j
  v_j(\lambda)}{\sigma_k} \right]^2
\end{eqnarray}
over the $a_j$, where $\sigma_k$ is the standard deviation of the
error in band $k$. Since we will use $v_j(\lambda)$ which are the light
emitted from stars and ionized gas, we make the reasonable assumption
that these only {\it add} light to the spectrum, not subtract it, and
require nonnegativity: that $a_j\ge 0$ for all $j$. (We discuss dust
below).  In order to perform this nonnegative minimization of Equation
\ref{chi2}, we use the method of \citet{sha02a}.

At this point, the reader probably realizes that the number of
templates $j$ should not be much higher than the number of bands $k$,
or else the problem becomes very degenerate. In fact, for the case of
unconstrained minimization of Equation \ref{chi2} the problem is
formally degenerate if $j>k$. The nonnegativity constraint adds enough
information that this is no longer the case, but still one cannot get
strong constraints on a huge number of parameters given a small number
of measurements. Nevertheless, below we {\it will} use a huge number
of $j$ given the number of $k$, simply for the purposes of finding the
subspace of possible galaxy spectra which actually conform to galaxy
observations. We do not mean to imply that we can determine each of
the resulting parameters at high signal-to-noise.

For our model templates we use a set of spectra from the stellar
evolution synthesis results of PEGASE2 (\citealt{fioc97a}). We use 15
bursts of varying ages. The top panel of Figure \ref{sfh} shows the
star-formation history of each of these bursts. For each age we create
five templates with five different metallicities ($Z=0.03$, $0.02$,
$0.008$, $0.004$, and $0.001$). The bottom panel of Figure \ref{sfh}
shows the template for the $Z=0.008$ case at every other age, starting
with the oldest. The bottom panel of For each age and metallicity, we
create three templates with three different extinctions ($\tau_V=0$,
$1$, and $3$) using the Milky Way extinction law using the ``clumpy,''
``dusty'' model of \citet{witt00a}. This procedure results in 225
templates and spans a very large range of possible spectra. Our
procedure is designed to take this large dimensionality space and
determine a low dimensionality subspace in which galaxies reside.

We fit these models to a set of galaxies for which we know the
redshifts. The majority of the redshifts are from the SDSS, but a
small number of the redshifts are from CNOC2 ({\bf ref}). We consider
only redshifts with 1$^\circ$.25 of the J2000 Equator, between
$-60<\alpha<100$. We match the locations of these spectra to the SDSS
photometric catalog ({\bf describe}) in all of the observations in
this region and to the 2MASS Extended Source Catalog (XSC). In
passing, we note that there are large errors (many arcseconds) in the
catalog astrometry for CNOC2, and we needed to take care to account
for these errors. For spectra which had matches in multiple runs of
the SDSS, we take the (unweighted) mean of their photometry to reduce
errors {\bf how many is this done for?}. We select a random set of
these galaxies in such a way that the number per redshift is nearly
constant across redshift. For each galaxy, we consider its
extinction-corrected model photometry. Figure \ref{gmrz} shows the
observed $g-r$ colors and redshifts of the resulting set of
galaxies. Note that only for the lower redshift galaxies do we have
matches to the 2MASS XSC.

Figure \ref{kfit} shows the residuals between the observed and fit
colors after minimizing Equation \ref{chi2} for all the galaxies in
our sample. Interestingly, even this large set of parameters cannot
perfectly describe the trends of color with redshift. There are
deviations at the percent level or so in $ugriz$. The $J$-band 2MASS
observations at moderate redshift appear to be about $0.1$ magnitudes
off of the models; this discrepancy is probably related to the similar
discrepancies found by \citet{blanton03d} and \citet{baldry03a} when
comparing the luminosity density of the universe to models.  Our
attitude in this paper will be to trust the data in preference to the
models and so below we describe how we deal with these discrepancies.

Each many-parameter fit to the broad-band fluxes of each galaxy yields
a complete spectrum from which to derive $K$-corrections. It also
provides a value of $\chi^2$; in principle, for galaxies without
redshifts we can find the assumed redshift with minimum $\chi^2$ in
order to find the photometric redshift. In practice, the fit takes too
long to perform to be useful for either of these purposes. Thus, we
describe here our method for reducing the dimensionality to a set of
three templates which adequately describe the data.

Our procedure has two steps. First, we use principle component
analysis on the set of 225-coefficent vectors describing all the fits
in order to find a 10-dimensional subspace which still describes the
data. Second, within this 10-dimensional subspace we group the
galaxies into a set of ten groups, average the spectra within each
group, and pick the triple among these groups which best describes the
data. We adopted this procedure because we did not know of a
principled way in which to reduce the dimensionality without
sacrificing nonnegativity; however, it turns out that such methods do
exist, and we recommend those methods for future investigators ({\bf
refs from sam}) while retaining this method for the moment because it
appears to work for our purposes.

Because the galaxies all have very different bolometric fluxes, there
is a strong variation in the coefficient values simply due to the
total flux differing by large factors. Therefore, before performing
principal component analysis, we scale all the coefficients of each
galaxy by a single number (per galaxy) such that the galaxies all lie
exactly on a plane in coefficient space. We choose the plane to be the
plane on which the sum of the templates has a constant flux between
1500 \AA\ and 30000 \AA. After scaling the galaxies to this plane, we
perform principal component analysis to obtain the five most principal
axes, using the efficient method of \citet{roweis99a}. There is almost
no difference between the full fit shown above and that using the
first five eigencoefficients.

On the other hand, the five eigencomponents natural consist partly of
positive star-formation components and partly of negative
star-formation components. This means that we cannot simply sum them
nonnegatively to get a nonnegative star-formation history. In order to
get a set of components that consists purely of all-positive
star-formation histories, we consider the distribution of galaxies in
the five-component space. We perform a $K$-means grouping on this set
of points, which produces ten groups. This is simply a way of very
roughly mapping the locus on which the galaxies reside (it is
significantly nonlinear within the five-dimensional space, so we
cannot simply take fewer principal components). We then average the
spectra in each group. This gives ten spectra who roughly span the
space of galaxy types. Finally, we pick from among those ten the
triple which best describes the galaxies, in the sense of minimizing
$\chi^2$ in Equation \ref{chi2}.

Figure \ref{kfit3} shows the color residuals after fitting with the
final three templates.

[ new model: 3 PEGASE2 spectra; Figure show spectra, Figure stellar
  mass to light ratios ]

[ to deal with that for K-correction and photo-z, tweak ]

We recognize that this procedure is rather convoluted and a bit of a
hack. As noted above, we have adopted it simply as one way of
automatically getting a small set of templates that explain galaxy
broad-band fluxes as a sum of star-formation histories. There are
methods for directly performing principal component analysis subject
to the conditions that the components and the coefficients are all
positive, and we recommend the investigation of such methods to the
reader ({\bf ref}).

\section{Determining $K$-corrections} 
\label{kcorrect}

Given a model spectrum for the galaxy, the determination of the
$K$-correction is straightforward. Here we repeat the

[ Figure demonstrate K-corrections ]

[ Figure show K-corrections to z=0 ]

[ note band-shifting ]

[ Figure test: against a bunch of spectra ]

\section{Determining photometric redshifts}
\label{kphotoz}

[ describe simple fitting procedure ]

[ Figure show training set results ]

[ Figure show results on deeper sdss data ]

\section{Public access to the code}

\section{Summary}
\label{conclusions}

[ summarize ]

[ this does not describe what constraints photometry puts on the SFH ]
